---
title: "Airline data in the United States from 2017 to 2023"
date: "2023/0/15"
output: html_notebook
---
Today, we make a representation from a specific airline/location about its passengers and the past 6 years.

```{r}
library(here)
library(nycflights13)
library(janitor)
library(tidyverse)
```

Now we load our files from the zip. This is from 2017 to May 2023. (https://www.transtats.bts.gov/DL_SelectFields.aspx?gnoyr_VQ=FIM&QO_fu146_anzr=Nv4+Pn44vr45).

```{r}
thisfile = here("raw", "DL_SelectFields_2017.zip")
df2017 = read_csv(thisfile) %>% clean_names()

thisfile = here("raw", "DL_SelectFields_2018.zip")
df2018 = read_csv(thisfile) %>% clean_names()

thisfile = here("raw", "DL_SelectFields_2019.zip")
df2019 = read_csv(thisfile) %>% clean_names()

thisfile = here("raw", "DL_SelectFields_2020.zip")
df2020 = read_csv(thisfile) %>% clean_names()

thisfile = here("raw", "DL_SelectFields_2021.zip")
df2021 = read_csv(thisfile) %>% clean_names()

thisfile = here("raw", "DL_SelectFields_2022.zip")
df2022 = read_csv(thisfile) %>% clean_names()

thisfile = here("raw", "DL_SelectFields_2023.zip")
df2023 = read_csv(thisfile) %>% clean_names()
```

Finally, we combine all the tables together and assign it to the variable df. Since we only care about the flights with passengers, we are going to filter that out.

```{r}
df = bind_rows(df2017, df2018, df2019, df2020, df2021, df2022, df2023)
df = df |> filter(passengers>0)
```

We want to graph the passengers and months to see which months of which year have what type of pattern.

```{r}
df2 = df |> filter(origin == "MIA" | dest == "MIA")
df2= df2 |> mutate(month_year = year+month/12)
ggplot(data = df2, mapping = aes(month_year)) + geom_histogram(aes(weight=passengers), color="black", fill="lightblue") + scale_x_continuous(breaks=2017:2023)
```
